---
title: "Final Report for Intro to Data Science"
output: html_notebook
---
# Team Members
* [Madison Yonash](mailto:myonash8902@floridapoly.edu)

# Introduction


# *Tidying* Up the Data

Since the data initially came from all years that I have used the app, it is important to make the data tidy and easier to work with.

Load necessary packages: 
```{r}
library(tidyverse)
library(lubridate)
```

Import data, which can be found [here](https://github.com/madiyonash/DS_Final_Project/blob/master/data/clean_report.csv):
```{r}
clean_report <- read_csv("https://raw.githubusercontent.com/madiyonash/DS_Final_Project/master/data/clean_report.csv?token=GHSAT0AAAAAABTIKYNMS4WBFEFLSOG2TB2SYTHDGIQ")
```

To analyze the data from 2019, we must filter the dataset for entries taken during that year. The use of the `lubridate` package makes this easy: 
```{r}
report_2019 <- clean_report %>% 
  filter(between(date, as.Date("2019-01-01"), as.Date("2019-12-31")))
```

If you take a glance at the `activities` column, you may notice that the observations have a varying number of activities. Here is a sample of some of the observations:
```{r}
report_2019 %>% 
  select(activities) %>% 
  sample_n(5)
```

To make it easier to work with this data, we can separate the activities so that each activity is contained within its own observation:
```{r}
report_2019_sep <- report_2019 %>% 
  separate_rows(activities, sep = "\\|")

#print sample of this report
report_2019_sep %>% 
  select(activities) %>% 
  sample_n(5)
```

Additionally, we can create an additional column that summarizes the average the mood per day. This will let us compare the rating of any particular day in 2019 with other factors.
```{r}
report_2019_day <- report_2019 %>% 
  group_by(date) %>% 
  summarize(ave_mood = mean(mood_scale, na.rm=T))
glimpse(report_2019_day)
```


# Data Description

# Exploratory Data Analysis

## Univariate Data Analysis

## Bivariate Data Analysis

## Data Analysis

# Conclusion

# Citation of Data

